Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
check
added reason to first bisim rule

\newcommand{\ineffCEK}{$\mathcal{X}$}
\newcommand{\ineffCEKD}{$\mathcal{D}$}
\newcommand{\ineffCEKUD}{$\mathcal{UD}$}

@@

\newcommand{\effCEK}[1]{$\mathcal{S}(#1)$}

@@

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation.  It is
the smallest congruence relation that also includes the pairs induced
by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending them with 
the same cast gives related continuations.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a naked function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both sides.
%
Rules for products and sums are similar.

\begin{figure}
  $\text{\effCEK{C}} \approx \text{\ineffCEKD}$
% \[\dots\]

  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : I
  }{
    \hcvOOinj{I}{v_1} \approx \vOOcast{v_2}{\cOOcast{I}{l}{\TOOdyn}}
    : \TOOdyn
  }
\\ \\
   \inference{}{
    \hcvOOtt \approx \vOOtt : \POOunit
   }
   \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\ \\
  \inference{
    \mathcal{E}_1 \approx \mathcal{E}_2 &
    dom(\mathcal{E}_1) = dom(\mathcal{E}_2) = \Gamma \\
    \Gamma , x : T_1 \vdash e : T_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\ \\
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\ \\
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2} \\
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
\\ \\
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}

>>>>

\newcommand{\ineffCEK}{\ensuremath{\mathcal{X}}}
\newcommand{\ineffCEKD}{\ensuremath{\mathcal{D}}}
\newcommand{\ineffCEKUD}{\ensuremath{\mathcal{UD}}}

@@

\newcommand{\effCEK}[1]{\ensuremath{\mathcal{S}(#1)}}

@@

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation
$s_1 \approx s_2$, where $s_1 \in \effCEK{C}$ and $s_2 \in \ineffCEKD{}$.
It is the smallest congruence relation that also includes the
pairs induced by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation. This handles the way in which
\effCEK{C} uses $wrap$ to add identity coercions around continuations
whereas \ineffCEKD{} does not.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending them with 
the same cast gives related continuations.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a naked function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both sides.
%
Rules for products and sums are similar.

\begin{figure}
%  $\text{\effCEK{C}} \approx \text{\ineffCEKD}$
% \[\dots\]
  \fbox{$s \approx s$}
  % Because the grammar for states of the two machines are not identical,
  % I think it is worthwhile to write out these rules explicitly. -Jeremy
  \begin{gather*}
    \inference{dom(\mathcal{E}_1) = dom(\mathcal{E}_2) \\
               \forall x {:} T \in dom(\mathcal{E}_1),\,
                   \mathcal{E}_1(x) \approx \mathcal{E}_2(x) : T\\
               \kappa_1 \approx \kappa_2 : T}
              {\langle e, \mathcal{E}_1, \kappa_1 \rangle \approx
              \langle e, \mathcal{E}_2, \kappa_2 \rangle}
    %% \quad
    %% \inference{v_1 \approx v_2 : T \quad \kappa_1 \approx \kappa_2 : T}
    %%           {\langle v_1, \kappa_1 \rangle  \approx
    %%           \langle v_2, \kappa_2 \rangle}
    \quad
    \inference{}
              {\sOOhalt{o} \approx \sOOhalt{o}}
  \end{gather*}
  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : I
  }{
    \hcvOOinj{I}{v_1} \approx \vOOcast{v_2}{\cOOcast{I}{l}{\TOOdyn}}
    : \TOOdyn
  }
\\ \\
   \inference{}{
    \hcvOOtt \approx \vOOtt : \POOunit
   }
   \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\ \\
  \inference{
    \mathcal{E}_1 \approx \mathcal{E}_2 &
    dom(\mathcal{E}_1) = dom(\mathcal{E}_2) = \Gamma \\
    \Gamma , x : T_1 \vdash e : T_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\ \\
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\ \\
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2} \\
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
\\ \\
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}
